### 2016-10-04 Paper_v7.tex
Fixing spelling, adding some polish

\subsection*{Identification of a novel LEF-1 regulator}
During model identification, we found that consistent TGF-$\beta$ induced EMT from a stable epithelial cell population required an additional regulatory protein.
This protein, which we called hypothetical regulator 1 (YREG1), was required to mediate between SNAIL/SLUG transcriptional activity
and the upregulation of LEF-1 expression following TGF-$\beta$1/2 exposure.
SNAIL/SLUG are well known transcriptional repressors \citep{Hemavathy:2000aa,Hemavathy:2000ab,Dhasarathy:2011aa}, although there are a few studies
which suggest that at least SNAIL can also act as a transcriptional activator \citep{Guaita:2002aa}.
In the model, we assumed the expression of SNAIL/SLUG was likely regulated by AP1/SP1 \citep{Jackstadt:2013aa}.
Thus, upon receiving direct Snail/Slug and TGF-$\beta$3 signals, the model predicted enhanced SNAIL/SLUG expression, consistent with experimental observations.
TGF-$\beta$1/2 stimulation also induces LEF-1 expression. However, literature evidence suggested that LEF-1 expression was not strongly dependent upon AP1/SP1 activity \citep{Eastman:1999aa}.
Thus, either SNAIL/SLUG are acting as inducers (contrary to substantial biochemical evidence) or, they are repressing the expression of an intermediate repressor.
Given the biochemical evidence supporting SNAIL/SLUG as repressors, we created YREG1 a hypothetical intermediate repressor whose
expression is downregulated by SNAIL/SLUG. The literature data therefore suggested that YREG1 had two transcriptional targets, LEF-1 and TGF-$\beta$3.
By adding this regulator, our simulations became consistent with training and literature data.
Medici et al. suggested that feedback between $\beta$-catenin and LEF-1 was likely, although this feedback had yet to be identified \citep{Medici:2008fk}.
Low levels of YREG1 expression were present in all simulations to regulate the formation of the $\beta$-catenin-LEF-1 complex.
To test the effect of YREG1 on the epithelial population, we conducted over-expression and knockdown simulations on untreated cells (Fig. \ref{fg:F4}C and \ref{fg:F4}D).
In the absence of YREG1, most of the population failed to consistently to retain a stable epithelial state (Fig. \ref{fig:F4}D).
Conversely, YREG1 overexpression revealed an enhanced epithelial phenotype, while some inherently transformed cells moved towards a hybrid phenotype  (Fig. \ref{fig:F4}C).
Elevation of YREG1 repressed LEF-1 and TGF-$\beta$3 expression, thereby not allowing free $\beta$-catenin to form the $\beta$-catenin-LEF-1 complex which promotes mesenchymal gene expression, or SMAD activity following from autocrine TGF-$\beta$3 signaling. Taken together, we found that low YREG1 expression was necessary for producing a stabile the model with an epithelial phentoype that is inducable across multiple stimuli such as TGF-$\beta$1/2, TGF-$\beta$3 and SNAIl/SLUG transfection, as seen in the training objectives.

@@

\subsection*{TGF$-\beta$1/2 and VEGF-A exposure promotes phenotype heterogeneity through NFATc and phosphorylated Sp1}
While we captured the central tendency of many of the molecular features of EMT induction following TGF$-\beta$1/2 exposure, an often neglected but important emergent
feature of developmental and pathological programs is population heterogeneity \citep{Park:2010fu}.
We (and others) previously hypothesized that deterministic model ensembles can interrogate population behavior, at least at a course grained level \citep{Lequieu:2011fj}.
We tested this hypothesis by analyzing the response of the population of EMT models to extracellular cues and then comparing this response to flow cytometry studies.
We quantiied the phenotypic response of the individual members of the ensemble to TGF$-\beta$1/2 stimulation as a function of physical measurements of abundance of markers in system.
We evlautated these quantities for each member of the ensemble (N $\simeq$ 1400) for two downstream phenotypic markers, Vimentin (mesenchymal) and E-cadherin (epithelial)
following the addition of TGF$-\beta$1/2 alone (Fig. \ref{fg:F3}), and VEGF-A in combination with NFATc inhibitors (Fig. \ref{fg:F3}).
We were further able to use this analysis to examine the activity and effects of underlying signaling molecules.

We identified model subpopulations that exhibited different behaviors following exposure to TGF$-\beta$1/2 (Fig. \ref{fg:F3}A).
Analysis of the molecular signatures of these subpopulations suggested the abundance, localization and state of the Sp1, AP-1 and NFATc transcription factors controlled population heterogeneity.
The behavior of the majority of models ($>$80\%) respond to treatment in Fig \ref{fg:F3}A-C, moving off of the untreated population set (gray).
These models showed the classically expected behavior, a switch from an epithelial to mesenchymal phenotype following TGF$-\beta$1/2 exposure.
Models near resembling untreated cells  had evaluated nuclear localized phosphorylated Sp1, relative to non-induced cells.
Elevated Sp1 activity decreased E-cadherin expression through Slug-mediated inhibition, which in turn increased Vimentin expression through TGF$-\beta$3 autocrine signaling and the liberation of $\beta$-catenin.

However, the most biologically interesting behavior was exhibited by cells achieving a hybrid phenotype, most notable in a dual treatment condition (black arrow), but also present in a small percentage of untreated cells (Fig. \ref{fg:F3}A).
Models with this hybrid phenotype had elevated Sp1 and NFAT transcriptional activity, resulting in the simultaneous increase of \emph{both} Vimentin and E-cadherin expression (Fig. \ref{fg:F4}A).
Analysis of these hypothetical cells suggested they had \emph{abnormal} signaling; deregulated NFAT expression and nuclear localization promoted E-cadherin expression while TGF$-\beta$1/2 induced Sp1 action promoted Vimentin expression.

@@

\begin{figurey}
\includegraphics [width=1.0\linewidth] {./figs/Fig4_mechanism.pdf}
\caption{Analysis of underlying signaling responses.
(A) We examined the distribution of NFATc1 and AP1/SP1 in cells containing the hybrid phenotype (VEGF-A + TGF$-\beta$2 case), showing the potential for cells to express both SP1 and NFATc1 in a non exclusive manner.
(B) We were able to show a fit to an additional objective demonstrating the actiation of GSK3 through PI3K. Our model captures this activation through TGF$-\beta$3 signaling. LY294002 is a PI3K inhibitor.
(C) We identified a novel regulator of LEF1 called YREG1 that allows Snail/Slug to emulate an inducer by repressing YREG1, which was required to stabalize the untreated population. 
YREG1 overexpression revealed an enhanced epithelial phenotype, while some inherently transformed cells moved towards a hybrid phenotype.
(D) In the absence of YREG1, most of the population failed to consistently to retain a stable epithelial state.}\label{fg:F4}
\end{figure}

>>>>

\subsection*{Identification of a novel LEF-1 regulator}
During model identification, we found that consistent TGF-$\beta$ induced EMT from a stable epithelial cell population required an additional regulatory protein.
This protein, which we called hypothetical regulator 1 (YREG1), was required to mediate between SNAIL/SLUG transcriptional activity
and the upregulation of LEF-1 expression following TGF-$\beta$1/2 exposure.
SNAIL/SLUG are well known transcriptional repressors \citep{Hemavathy:2000aa,Hemavathy:2000ab,Dhasarathy:2011aa}, although there are a few studies
which suggest that at least SNAIL can also act as a transcriptional activator \citep{Guaita:2002aa}.
In the model, we assumed the expression of SNAIL/SLUG was likely regulated by AP1/SP1 \citep{Jackstadt:2013aa}.
Thus, upon receiving direct SNAIL/SLUG and TGF-$\beta$3 signals, the model predicted enhanced SNAIL/SLUG expression, consistent with experimental observations.
TGF-$\beta$1/2 stimulation also induces LEF-1 expression. However, literature evidence suggested that LEF-1 expression was not strongly dependent upon AP1/SP1 activity \citep{Eastman:1999aa}.
Thus, either SNAIL/SLUG are acting as inducers (contrary to substantial biochemical evidence) or, they are repressing the expression of an intermediate repressor.
Given the biochemical evidence supporting SNAIL/SLUG as repressors, we created the hypothetical YREG1 repressor whose
expression is downregulated by SNAIL/SLUG. The literature data therefore suggested that YREG1 had two transcriptional targets, LEF-1 and TGF-$\beta$3.
By adding this regulator, our simulations became consistent with training and literature data.
Medici et al. suggested that feedback between $\beta$-catenin and LEF-1 was likely, although this feedback had yet to be identified \citep{Medici:2008fk}.
Low levels of YREG1 expression were present in all simulations to regulate the formation of the $\beta$-catenin-LEF-1 complex.
To test the effect of YREG1 on the epithelial population, we conducted over-expression and knockdown simulations on untreated cells (Fig. \ref{fg:F4}C and \ref{fg:F4}D).
In the absence of YREG1, the population of models failed to consistently to retain a stable epithelial state (Fig. \ref{fg:F4}D).
Conversely, YREG1 amplification revealed an enhanced epithelial phenotype, while some inherently transformed cells moved towards a hybrid phenotype  (Fig. \ref{fg:F4}C).
Elevated YREG1 repressed LEF-1 and TGF-$\beta$3 expression, thereby not allowing free $\beta$-catenin to form the $\beta$-catenin-LEF-1 complex, or TGF-$\beta$3 induced SMAD activation.
Taken together, low YREG1 expression was required for the maintenance of a stable epithelial phenotype that was simultaneously inducible across TGF-$\beta$1/2, TGF-$\beta$3 and SNAIl/SLUG transfection, as seen in the training objectives.

@@

\subsection*{TGF$-\beta$1/2 and VEGF-A exposure promotes phenotype heterogeneity through NFATc and phosphorylated Sp1}
While we captured the central tendency of many of the molecular features of EMT induction following TGF$-\beta$1/2 exposure, an often neglected but important emergent
feature of developmental and pathological programs is population heterogeneity \citep{Park:2010fu}.
We (and others) have previously hypothesized that deterministic model ensembles can simulate population behavior, at least at a course grained level \citep{Lequieu:2011fj}.
We tested this hypothesis by analyzing the response of the population of EMT models to extracellular cues and then comparing this response to flow cytometry studies.
We quantified the phenotypic response of the individual members of the ensemble to TGF$-\beta$1/2 stimulation for two downstream phenotypic markers, Vimentin (mesenchymal) and E-cadherin (epithelial) following the addition of TGF$-\beta$1/2 alone (Fig. \ref{fg:F3}), and VEGF-A in combination with NFATc inhibitors (Fig. \ref{fg:F3}).
%We were further able to use this analysis to examine the activity and effects of underlying signaling molecules.

We identified model subpopulations that exhibited different behaviors following exposure to TGF$-\beta$1/2 (Fig. \ref{fg:F3}A).
Analysis of the molecular signatures of these subpopulations suggested the abundance, localization and state of the Sp1, AP-1 and NFATc transcription factors controlled population heterogeneity.
The majority of models ($>$80\%) responded to treatment, moving off of the untreated population (Fig \ref{fg:F3}A-C, gray).
These models showed the classically expected behavior, a switch from an epithelial to mesenchymal phenotype following TGF$-\beta$1/2 exposure.
Models near resembling untreated cells had evaluated nuclear localized phosphorylated Sp1, relative to non-induced cells.
Elevated Sp1 activity decreased E-cadherin expression through Slug-mediated inhibition, which in turn increased Vimentin expression through TGF$-\beta$3 autocrine signaling and the liberation of $\beta$-catenin. However, the most biologically interesting behavior was exhibited by cells achieving a hybrid phenotype, most notable in a dual treatment condition (black arrow), but also present in a small percentage of untreated cells (Fig. \ref{fg:F3}A).
Models with this hybrid phenotype had elevated Sp1 and NFAT transcriptional activity, resulting in the simultaneous increase of \emph{both} Vimentin and E-cadherin expression (Fig. \ref{fg:F4}A).
Analysis of these hypothetical cells suggested they had deregulated NFAT expression and nuclear localization promoted E-cadherin expression while TGF$-\beta$1/2 induced Sp1 action promoted Vimentin expression.

@@

\begin{figure}
\includegraphics [width=1.0\linewidth] {./figs/Fig4_mechanism.pdf}
\caption{Analysis of underlying signaling responses.
(A) We examined the distribution of NFATc1 and AP1/SP1 in cells containing the hybrid phenotype (VEGF-A + TGF$-\beta$2 case), showing the potential for cells to express both SP1 and NFATc1 in a non exclusive manner.
(B) We were able to show a fit to an additional objective demonstrating the actiation of GSK3 through PI3K. Our model captures this activation through TGF$-\beta$3 signaling. LY294002 is a PI3K inhibitor.
(C) We identified a novel regulator of LEF1 called YREG1 that allows Snail/Slug to emulate an inducer by repressing YREG1, which was required to stabalize the untreated population.
YREG1 overexpression revealed an enhanced epithelial phenotype, while some inherently transformed cells moved towards a hybrid phenotype.
(D) In the absence of YREG1, most of the population failed to consistently to retain a stable epithelial state.}\label{fg:F4}
\end{figure}
